---
title: "eds-223-week1-discussion"
format: html
---
# Part 1
Load the neccessary packages. 
```{r}
library(here)
library(tidyverse)
library(sf)
library(stars)
library(tmap)
```

Next, read in the necessary files. 
```{r}
ei_points <- st_read(here("data", "easter_island", "ei_points.gpkg")) |> 
  filter(type == "volcano")
ei_elev <- read_stars(here("data", "easter_island", "ei_elev.tif"))
ei_borders <- st_read(here("data", "easter_island", "ei_border.gpkg"))
ei_roads <- st_read(here("data", "easter_island", "ei_roads.gpkg"))
```

## Creating a single-object map 
Use tm_shape() + tm_borders() to create a map outline of Rapa Nui.
```{r}
tm_shape(ei_borders) + tm_borders()
```

## Create a multiple-object map with tmap
Expand on your single-object map to create a map of Rapa Nui and…

…denote the island’s borders and continuous elevation
…denote the island’s volcanoes and roads
…play with the color palette and essential map elements
```{r}
#palette_colors = c("#470C00", "#751400", "#A31B00", "#FF512E", )
#ei_elev, palette =
tm_shape(ei_borders) + tm_borders() + tm_shape(ei_elev) + 
  #tm_layout() 
   tm_raster(col.scale  = tm_scale(values = c("darkgreen", "lightgreen")), col.legend = tm_legend(title = "Elevation (m)")) + tm_shape(ei_roads) + tm_layout() + tm_lines() + 
  tm_shape(ei_points) + tm_symbols(shape = 24, size = .7, fill = "red")+ tm_text("name", size = .7, col = "black") + tm_layout(outer.bg.color = "beige")

```

#Create an interactive map with tmap
Let's make this interactive with tmap_view(). 
```{r}
tmap_mode("plot")

tm_shape(ei_borders) + tm_borders() + tm_shape(ei_elev) + 
  #tm_layout() 
   tm_raster(col.scale  = tm_scale(values = c("brown", "red")), col.legend = tm_legend(title = "Elevation (m)")) + tm_shape(ei_roads) + tm_layout() + tm_lines() + 
  tm_shape(ei_points) +
  #tm_polygons() +
  tm_symbols(shape = 24, size = "sv", col = "black", fill = "orange", fill.legend = tm_legend(title = "Volcano Classification")) + 
  tm_text("name", size = .7, col = "white") + tm_layout(outer.bg.color = "beige") 


tm_shape(ei_borders) + tm_borders() + tm_shape(ei_elev) + 
  #tm_layout() 
   tm_raster(col.scale  = tm_scale(values = c("darkgreen", "lightgreen")), col.legend = tm_legend(title = "Elevation (m)")) + tm_shape(ei_roads) + tm_layout() + tm_lines() + 
  tm_shape(ei_points) + tm_symbols(shape = 24, size = .7, fill = "red")+ tm_text("name", size = .7, col = "black") + tm_layout(outer.bg.color = "beige", bg.color = "lightblue")




```

## Using ggplot to create a map
```{r}
ggplot(data = ei_borders) + geom_sf() 
```


